---
title: "R-programming Project - Analysis and Prediction of Airbnb Listing Prices"
author: "Nived Suresan"
date: "2023-07-07"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

The goal of this project is to perform an exploratory data analysis (EDA) and predictive modelling on Airbnb listing data using the R programming language. The project will encompass all stages of the data science lifecycle, from data import and cleaning to visualisation and modelling. 

## Data Loading
```{r}
library(dplyr)
library(caret)
library(tinytex)

#read the data
data = read.csv("C:/Users/Nived/Downloads/listings.csv~1/listings1.csv")

```

## Data Tidying

```{r}
#2. removing unwanted columns from the data for analysis
data = data[,c('neighbourhood','room_type','latitude','longitude','accommodates','bedrooms','beds','price','review_scores_rating','reviews_per_month')]

#3. converting the price column which is a string into an integer
data$price = gsub("\\$","",data$price)
data$price <- as.integer(data$price)
```

## Data Exploration

```{r}
#summarize the data
summary(data)
```

## Feature engineering

```{r}
data = data %>% 
  mutate(price_per_person = price %/% accommodates) %>%
  mutate(beds_per_room = beds %/% bedrooms) %>%
  mutate(beds_group = case_when(
    beds_per_room == 1 ~ "Single bed",
    beds_per_room == 2 ~ "Double bed",
    beds_per_room >= 3 ~ "Multiple beds"
  ))

#finding the average price paid by the customer based the number of people staying
avg = aggregate(data$price ~ data$accommodates, data=data,FUN=mean)

room_type_freq = table(data$room_type)
room_type_freq = data.frame(room_type_freq)

#removes all rows with missing values and renaming the column names
avg = na.omit(avg)
colnames(avg) = c("accommodates","price")

```

## Bar Plot

```{r}
barplot(room_type_freq$Freq, names.arg=room_type_freq$Var1,xlab="Room Type",ylab="Frequency")
```

## Scatter Plot and Regression
```{r}

plot(avg$accommodates, avg$price, xlim=c(1,16),ylim=c(0,800), xlab ="accommodate", ylab="price")

#Splitting the data into training set and testing set
set.seed(123)
train_index <- createDataPartition(avg$price, p = 0.7, list = FALSE)
train_data <- avg[train_index, ]
test_data <- avg[-train_index, ]

# Build the linear regression model
model <- lm(price ~ accommodates, data = train_data)

# Make predictions on the test set
predictions <- predict(model, newdata = test_data)
rmse <- sqrt(mean((test_data$price - predictions)^2))
print(rmse)

abline(model,col='blue')

summary(model)


```

## Conclusion

The overall p-value, labeled as "Pr(>|t|)" or "Pr(>F)" in the summary output, is 0.000177. This p-value is less than 0.05, indicating that the regression model is significant. Typically, If the p-value is less than 5% or 0.05, then the model is considered significant.

Moreover, the adjusted R-squared value of 0.6789 suggests that approximately 67.89% of the variance in the dependent variable in the model.This means that the price of stay can be estimated using the accommodates.

Later we evaluate the RMSE value of the data set by splitting into training set and testing
set. We then model the training set data using linear regression and predict based on this model. after finding the rmse value we get a value 30.5345 which tells us about the error of the model.This is sufficienty accurate when It comes to predicting prices.

